Dark Control: The Default Mode Network as a Reinforcement Learning Agent

### 2018-08-08 paper/paper.tex
some edits of edits

The present work adopts the perspective of
a human \textit{agent} faced with the choice of the next actions
and guided by outcomes to optimize behavioral performance. These outcomes
can be experienced, hypothetically imagined, or simply future expectations.
Formally, we propose reinforcement learning as a particularly attractive framework
for describing and quantifying the DMN function.
An intelligent agent improves the interaction with the environment
by continuously updating its computation of value estimates and action predispositions
through integration of feedback outcomes.
That is, ``[agents], with their actions, modify the environment and in doing so
partially determine their next stimuli, in particular stimuli that are necessary
for triggering the next action''
\citep{pezzulo2011grounding}.
Agents with other behavioral policies therefore sample different
distributions of action-perception trajectories \citep{ghavamzadeh2015bayesian}.
Henceforth,
\textit{control} refers to the influence that an agent exerts when interacting
with the environment to reach preferred states.

@@

  \subsubsection{Self-training and the loss function}
  \label{sec:self}
Successful learning in brains and computer algorithms may not be possible without a
defined learning goal \textemdash the \textit{loss function}.
The action $a$ chosen in state $s$ according to the policy matrix defined in
\eqref{eq:policy} yields a reward $r$ collected by the agent,
% (measured on their internal currency of worldly values ?),
after which the environment transitions to a new state $s' \in \mathcal S$.
% \textbf{XXX Please try to smoothen the previous sentence}
One such cycle yields a new \textit{experience} $e = (s,a,r,s')$.
Each cycle represents a behavior unit of the agent
and is recorded in replay memory buffer \textemdash
which we hypothesize to be subserved by the
HC \textemdash, possibly discarding the oldest entries to make space:
$\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1}
+ \delta \theta_{k}$ of the parameters for its approximate model of the
$Q$-value function. This warrants a learning process and definition of a loss
function. % \subsubsection{Stochastic gradient in the DMN via experience replay}
The Bellman equation \eqref{eq:bellman} provides a way to obtain such a loss
function \eqref{eq:oracle} as we outline in the following.
% A practically successful RL approach
% needs ways to efficiently approximate the state-action value function
% to allow for scaling to large state and action spaces.
Experience replay consists in sampling
% (uniform or importance-weighted
% \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.})
batches of experiences $e$
$(s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$.
The agent then tries to approximate
the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the
Bellman equation \eqref{eq:bellman}, namely
\begin{equation}
  y_k := y_k(s,a,s') =  r + \gamma \max_{a'} \tilde{Q}(s', a'|\theta_{k-1}),
\end{equation}
with the estimation of a parametrized regression model $(s,a)
\mapsto \tilde{Q}(s, a|\theta_{k-1})$.
From a neurobiological perspective, experience replay can be manifested as the
re-occurrence of neuron spiking sequences that have also been measured during
specific prior actions and environmental states. The HC is a strong candidate
for contributing to such neural reinstantiation of behavioral episodes as
neuroscience experiments have repeatedly indicated in rats, mice, cats, rabbits,
songbirds, and monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}. Refer to
Fig. \ref{fig:rl_process_chart}.

>>>>

The present work adopts the perspective of
a human \textit{agent} faced with the choice of the next actions
and guided by outcomes to optimize behavioral performance. These outcomes
can be really experienced, hypothetically imagined, or simply expected in the future.
Formally, we propose reinforcement learning as a particularly attractive framework
for describing, containing, and quantifying the DMN function.
An intelligent agent improves the interaction with the environment
by continuously updating its computation of value estimates and action predispositions
through integration of feedback outcomes.
That is, ``[agents], with their actions, modify the environment and in doing so
partially determine their next stimuli, in particular stimuli that are necessary
for triggering the next action''
\citep{pezzulo2011grounding}.
Agents with other behavioral policies therefore sample different
distributions of action-perception trajectories \citep{ghavamzadeh2015bayesian}.
Henceforth,
\textit{control} refers to the influence that an agent exerts when interacting
with the environment to reach preferred states.

@@

  \subsubsection{Self-training and the loss function}
  \label{sec:self}
Successful learning in brains and computer algorithms may not be possible without a
defined learning goal \textemdash the \textit{loss function}.
The action $a$ chosen in state $s$ according to the policy matrix defined in
\eqref{eq:policy} yields a reward $r$ collected by the agent,
% (measured on their internal currency of worldly values ?),
after which the environment transitions to a new state $s' \in \mathcal S$.
% \textbf{XXX Please try to smoothen the previous sentence}
One such cycle yields a new \textit{experience} $e = (s,a,r,s')$.
Each cycle represents a behavior unit of the agent
and is recorded in replay memory buffer \textemdash
which we hypothesize to be subserved by the
HC \textemdash, possibly discarding the oldest entries to make space:
$\mathcal D \leftarrow \text{append}(\mathcal D, e)$.
At time step $k$, the agent seeks an update $\theta_{k} \leftarrow \theta_{k-1}
+ \delta \theta_{k}$ of the parameters for its approximate model of the
$Q$-value function. Step-by-step model parameter updates warrant a learning process and definition of a loss
function. % \subsubsection{Stochastic gradient in the DMN via experience replay}
The Bellman equation \eqref{eq:bellman} provides a way to obtain such a loss
function \eqref{eq:oracle} as we outline in the following.
% A practically successful RL approach
% needs ways to efficiently approximate the state-action value function
% to allow for scaling to large state and action spaces.
Experience replay consists in sampling
% (uniform or importance-weighted
% \footnote{e.g weighted by TD error of the state transition $s \overset{a}{\rightarrow} s'$.})
batches of experiences $e$
$(s, a, r, s') \sim \mathcal D$ from the replay memory $\mathcal D$.
The agent then tries to approximate
the would-be $Q$-value for the state-action pair $(s,a)$ as predicted by the
Bellman equation \eqref{eq:bellman}, namely
\begin{equation}
  y_k := y_k(s,a,s') =  r + \gamma \max_{a'} \tilde{Q}(s', a'|\theta_{k-1}),
\end{equation}
with the estimation of a parametrized regression model $(s,a)
\mapsto \tilde{Q}(s, a|\theta_{k-1})$.
From a neurobiological perspective, experience replay can be manifested as the
re-occurrence of neuron spiking sequences that have also been measured during
specific prior actions or environmental states. The HC is a strong candidate
for contributing to such neural reinstantiation of behavioral episodes as
neuroscience experiments have repeatedly indicated in rats, mice, cats, rabbits,
songbirds, and monkeys \citep{buhry2011,nokia2010,dave2000,skaggs2007}. Refer to
Fig. \ref{fig:rl_process_chart}.
